\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2020-06-17 redaction/submission_JEGeo/revision/revised_article/DHP_JoEG_rev1.tex
modif prg FS + section endo
Modification prog first stage pour avoir tableau de stat desc en latex
Premier draft de la sous-section "Endogeneity concerns", pour le moment placée en fin de section 2

\paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden.
Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price). Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair.

It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components. However, the latter can be biased by the abovementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$.

XXX propose firss stage equation, explain IVs XXX ICI XXX

XXX provide results 2nd stage

>>>>

\paragraph{Endogeneity concerns} One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to an endogeneity bias, as the price set by the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the FAS price). 
Besides, related concerns may arise from productivity-sorting model in \citet{melitz} or the quality-sorting model in \citet{baldwin_harrigan}: more productive firms and/or firms selling high-quality products will charge higher prices, all other things equal – in our case, for a given country-product pair. In addition, there is also the possibility that bigger firms might impact transport costs, due to their ability of bargaining discounts for larger shipped volumes.

It is true we are not interested in causal inference, since our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components. However, the latter can be biased by the aforementioned endogeneity issues since they imply that $\frac{1}{\widetilde{p}_{ik}}$ is correlated in one direction on another with residuals $\epsilon_{ik}$. This may ultimately harm the estimation of $\widehat{t}/\widetilde{p}$ and consequently, $\widehat{\tau}$.

Therefore, we propose to compare our main estimates with those stemming from a two-stage least-squares procedure. Put differently, we will replace in the right-hand side of Equation (\ref{eq:model_IetA}) $\widetilde{p}$ by predicted values stemming from the following equation:

\begin{equation}
\ln\left(\widetilde{p}_{ik}\right)= \beta IV_{ik} +\lambda_{is(k)}+ \mu_{ik} \label{eq:first_stage}
\end{equation}

\noindent where $\lambda_{is(k)}$ represents country-3-digit sector fixed effects, and $IV_{ik} $ the instrumental variable used to purge $\widetilde{p}_{ik}$ from endogeneity issues. Following common practice in the literature, we will use as instrument 5-digit product-specific tariff rates, or alternatively, a combination between the latter and lagged fas prices (XXX TO BE DONE XXX). Consistently with our overall estimation strategy, Equation (\ref{eq:first_stage}) is estimated on a yearly basis.

Tables \ref{tab: FS_air} and \ref{tab: FS_ves} report summary results of these first stage regressions:

\begin{table}[htbp]\centering
\caption{Summary Results for First-Stage Regressions - Air\label{tab: FS_air}}
\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
\textbf{variable } & \textbf{mean} & \textbf{p25} & \textbf{p50} & \textbf{p75} & \textbf{sd} & \textbf{min} & \textbf{max} \\
\hline
   $\beta_{IV}$ &    -0.0858 &    -0.1247 &    -0.0944 &    -0.0362 &     0.0469 &    -0.1741 &    -0.0128 \\
        $sd_{\beta}$  &     0.0184 &     0.0139 &     0.0197 &     0.0237 &     0.0082 &     0.0048 &     0.0331 \\
  Student t  &    -5.2255 &    -6.5428 &    -5.2703 &    -3.7943 &     2.4782 &   -11.4557 &    -0.4184 \\
    F stat  &    33.2938 &    14.4215 &    27.7801 &    42.8707 &    28.1280 &     0.1751 &   131.2323 \\
  adj R²  &     0.4845 &     0.4712 &     0.4768 &     0.4934 &     0.0227 &     0.4452 &     0.5472 \\
  R² within &     0.0025 &     0.0015 &     0.0023 &     0.0036 &     0.0016 &     0.0000 &     0.0064 \\
\hline
\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} }}
\end{tabular}
\end{table}

\begin{table}[htbp]\centering
\caption{Summary Results for First-Stage Regressions -Ves\label{tab: FS_ves} }
\begin{tabular} {@{} l r r r r r r r @{}} \\ \hline
\textbf{variable } & \textbf{      mean} & \textbf{       p25} & \textbf{ p50} & \textbf{p75} & \textbf{sd} & \textbf{       min} & \textbf{       max} \\
\hline
   $\beta_{IV}$ &     0.0584 &     0.0180 &     0.0733 &     0.0846 &     0.0349 &    -0.0010 &     0.1147 \\
       $sd_{\beta}$  &     0.0124 &     0.0108 &     0.0138 &     0.0155 &     0.0047 &     0.0041 &     0.0195 \\
  Student t  &     4.5161 &     3.3592 &     4.8718 &     5.5721 &     1.9626 &    -0.0781 &     8.9563 \\
    F stat  &    24.1509 &    11.4029 &    23.7486 &    31.0489 &    17.2397 &     0.0061 &    80.2158 \\
  adj R² &     0.5856 &     0.5767 &     0.5878 &     0.5960 &     0.0172 &     0.5364 &     0.6323 \\
  R² within &     0.0020 &     0.0006 &     0.0021 &     0.0032 &     0.0015 &     0.0000 &     0.0062 \\
\hline
\multicolumn{8}{@{}l}{\footnotesize{\emph{Source:} }}
\end{tabular}
\end{table}

Both Tables suggest that custom duties are a very good predictor of fas prices $\widetilde{p}_{ik}$, with strong significance. The F-statistic for significance of the instrument in the first-stage is on average well above 10, the ``rule-of-thumb'' threshold value recommended by \citet{staiger_stock} as a test for weak instrumentation.

XXX on peut aussi ajouter des graphs des beta estimés ici, peut-être éventuellement avec les IC XXX

XXX Reste ici l'épineux problème de la différence de signe entre les deux modes de transport; avant de creuser, voir si on ne résoudrait pas le problème avec l'IV combinant duties et lagged prices XXX

Provided that the instrument used in Equation \ref{eq:first_stage} displays the right statistical properties, we can confidently re-inject the predictions it produces for $\widetilde{p}_{ik}$ on the right-hand side of Equation (\ref{eq:model_IetA}\footnote{Note that, since the predictions are expressed in logarithms, they are transformed through the exponential function before their implementation in Equation (\ref{eq:model_IetA}.} to produce a 2SLS-type of estimation. Results are displayed below, as well as a measure of the gap with our main estimates.

XXX We expect the gap to be very small XXX
